### 2013-01-31 drafts/ijcai2013/ijcai13.tex
Changes.

\documentclass{article}

\usepackage{ijcai13}
\usepackage{times}
\usepackage{amsmath} 
\usepackage{latexsym} 
\usepackage{ dsfont }
\usepackage{tikz}
\usepackage{tikz-qtree}
\usepackage{float}
\usepackage{circuitikz}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{alltt}
\usepackage{dashrule}
\usepackage{array}
\usepackage[ruled,vlined,linesnumbered]{algorithm2e}
\usetikzlibrary{arrows,automata, positioning, patterns,backgrounds}

\newtheorem{definition}{Definition}
\DeclareMathOperator*{\argmax}{arg\,max} 
\DeclareMathOperator*{\argmin}{arg\,min}

\title{Bootstrap learning via modular concept discovery}

\author{Eyal Dechter \\
MIT\\
USA \\
edechter@mit.edu
\And
Jon Malmaud \\
MIT \\ 
USA \\
malmaud@mit.edu
\And 
Ryan P.  Adams \\
Harvard University\\
USA \\
rpa@seas.harvard.edu
\And
Joshua B. Tenenbaum \\
MIT\\
USA \\
jbt@mit.edu
}

\begin{document}

\maketitle

\begin{abstract}
 Suppose a learner is faced with a domain of problems about which it
 knows nearly nothing. It does not know the distribution of problems,
 the space of solutions is not smooth, and the reward signal is
 uninformative, providing perhaps a few bits of information but not
 enough to steer it effectively. How can such a learner ever get off
 the ground? A common intuition is that if the solutions to these
 problems share a common structure, and the learner can solve some
 simple problems by brute force, it should be able to extract useful
 components from these solutions and, by composing them, explore the
 solution space more efficiently. Here, we formalize this intuition,
 where the solution space is the space of typed functional programs
 and the gained information is stored as a stochastic grammar over
 programs. We propose an iterative procedure for exploring such
 spaces: in the first step of each iteration, the learner explores a
 finite subset of the domain, guided by a stochastic grammar; in the
 second step, the learner compresses the successful solutions from the
 first step to estimate a new stochastic grammar. We test this
 procedure on symbolic regression and boolean circuit learning and
 show that the learner discovers modular concepts for these
 domains. Whereas the learner is able to solve almost none of the
 posed problems in the procedure's first iteration, it rapidly becomes
 able to solve a large number by gaining abstract knowledge of the
 structure of the solution space.

\end{abstract}

\section{Introduction}
Many of the problems we want to solve in AI are best cast as search
problems. There are many search algorithms, but they typically exploit
local or topological structure in the search space. For example,
convex optimization and gradient based search methods~\cite{} take
advantage of local smoothness to move in good directions. Local search
algorithms move from one candidate solution to a neighboring
one\cite{}. Heuristic search techniques take advantage of topological
structure to prune off large parts of the solution space~\cite{}.

However, many of the problems we want solve -- discovering scientific
theories, developing new technologies, figuring out how to bake -- do
not come with such a structured search space. This is particularly
true of problems, as these are, whose solutions can be naturally
framed as programs. Here we ask how to extract structure from the
space of programs by attempting to solve multiple related problems in
a given domain at once.

We take as inspiration the programmer who -- having reused several
motifs, programming patterns, or subroutines to solve different but
similar simple tasks -- creates a function library to help manage the
complexity of larger projects. Our algorithm builds a library
of reusable program components and a distribution over these,
effectively learning from simple tasks a search space structure which
enables it to solve more complex tasks.

In this work, we try to see how far this idea of reusable function
definition can get us. Our question is, Can an AI system discover the
structure latent in the solutions to multiple related problems and
thereby bootstrap effective exploration of the search space? Can
discovering modular program components transform a problem from one in
which search is intractable into a problem in which it becomes
increasingly feasible with experience?

Concretely, our algorithm works as follows. It begins with a small
library of program primitives and a distribution over programs
synthesized from this library. In each iteration, it generates
programs in order of highest probability according to the library
learned in the previous iteration. From the discovered solutions, it
finds new reusable functions that compress these solutions, and it
re-estimates the distribution over programs by weighting these new
functions according to their frequencies.

We show that this iterative approach allows us to achieve very high
performance in tasks where just searching according to the initial
distribution over program primitives clearly fails.

\section{Related Work}
The idea of discovering and using reusable subcomponents is an old one
in AI. Rendell's 1985 paper~\cite{DBLP:conf/ijcai/Rendell85},
``Substantial Constructive Induction Using Layered Information
Compression: Tractable Feature Formation in Search,'' -- which could
have been the name of this paper -- did not use programs as its
solution space representation. Within the Inductive Logic Programming
literature, researchers have explored ``Constructive Induction,''
generating new and reusable logic programming
terms~\cite{DBLP:conf/ijcai/Muggleton87}. We believe that applying
constructive induction to a compositional functional programming
paradigm will provide even greater benefits.

Dicovering useful subroutines has been explored in Genetic
Programming as
well~\cite{DBLP:conf/foga/Koza92}~\cite{DBLP:conf/ices/KozaBAK96}.

The importance of a curriculum has also been explored previously,
though generally the curriculum has been explicitly given rather than
discovered. Quinlan and Cameron-Jones~\cite{Quinlan93foil:a} showed
that a logic program learning algorithm could learn sequentially more
difficult list programs if the algorithm was provided with its own
solutions to previously solved simple problems. More recently, several
researchers have observed that presenting problems to machine learning
algorithms in order of difficulty results in better
performance~\cite{Krueger2009380}~\cite{DBLP:conf/icml/BengioLCW09}.

Our use of a stochastic grammar over combinatory logic is based on
Liang et al.~\cite{liang10programs}. In that work, a non-parametric
distribution was placed over combinatory logic expressions, and an
MCMC algorithm was developed to perform joint inference over these
expressions and the underlying stochastic grammar. Our approach is
inspired by this generative model, but avoids the difficulties
associated with a stochastic local search over programs (the same
difficulties faced by genetic programming), especially when the
likelihood function (or the fitness function) are not of much help.

%% AI and machine learning algorithms commonly take advantage of the
%% relationship between a performance metric (error, reward, fitness,
%% etc.) and the local or topological structure of the solution
%% space. Whether in discrete or continuous domains, the key is to design
%% such a relationship and write a algorithm that either moves in good
%% directions or prunes off large parts of the solution space. Whether or
%% not such a relationship exists depends on the representation, and much
%% work has gone into finding good representations for problem
%% solving. Recent work, especially in machine learning, has focused on
%% ``feature learning,'' with the explicit aim of discovered good
%% representational building blocks in vector-valued solutions spaces.

%% Many of the most natural solution representations for intelligent
%% agents, however, are not naturally expressed as vector-valued
%% spaces. Consider the problems of discovering scientific theories,
%% inventing new technologies, communicating with others, or executing a
%% complex motor routine (such as dancing or playing an instrument). The
%% solutions to these problems are most naturally expressed as programs,
%% that is, as the composition of modular components whose semantics are
%% determined by the actions of one component on another.

%% How can we learn good solution representations for these kinds of
%% tasks without building in the kind of domain knowledge that gives rise
%% to a useful relationship between representation and performance? How
%% can we get traction on a set of problems when small tweaks to
%% solutions give rise to wildly varying performances in an unpredictable
%% manner, or when, for most problems, most of the solutions fail in a
%% uniformly bad way?

%% A popular idea is that good representations are those that encode the
%% data space in an efficient way, that is, that enable compression of
%% the data~\cite{}. For program representations, this most naturally
%% corresponds to defining reusable components so that the total number
%% of unique expressions for a set of solutions is minimized (i.e., this
%% is directly akin to the subroutine libraries programmers use to manage
%% the complexity of their code).

%% In this work, we try to see how far reusable subroutine definition can
%% get us on its own. Our question is, Can discovering modular components
%% transform a problem from one in which brute-force search will not work
%% to a problem in which it does?

%% To emphasize an impoverished relationship between reward signals and
%% program representations, we make our tasks all-or-none: a solution
%% either solves the problem or it does not, and partial solutions do not
%% get any credit.

%% The intuitive approach to applying this idea is simple: if solutions
%% to problems in the domain share common structure, and there exist
%% problems whose solutions are simple enough to find by brute force,
%% then we should be able to extract a distribution over solution
%% subcomponents from these simple cases that allows us to run subsequent
%% searches more efficiently.

%% More concretely, our algorithm works as follows: each iteration, it
%% first generates programs in order of highest probability according to
%% a subroutine distribution learned in the previous iteration. From the
%% discovered solutions, it finds new reusable subroutines that compress
%% these solutions and estimates their probabilities.

%% We show that this iterative approach allows us to achieve very high
%% performance in tasks where just searching according to the initial
%% distribution over program primitives clearly fails.

%%  \hline

%% AI and machine learning algorithms commonly take advantage of the idea
%% that local structure in the solution space is related to local
%% structure in the outcome. In continuous domains, learning algorithms
%% algorithms use error gradients to guide search. In discrete domains,
%% whether in heuristic or local search paradigms, knowledge about the
%% solution space topology and its relation to reward and cost metrics
%% enables pruning, reasoning, and moving in the solution space more
%% efficiently. However, choosing the representations that have
%% beneficial structural properties requires a lot of general and domain
%% knowledge, and this what researchers do and what they cannot leave up
%% to their algorithms.

%% What is a good strategy for an agent when its representation of the
%% solution space is not smooth with respect to the outcome? Many hard
%% learning problems have this property, especially those in which the
%% solution space is most naturally represented as a set of
%% programs. Consider the problems of discovering scientific theories,
%% inventing new technologies, learning to cook, or learning to
%% communicate. These are problems in which 1) solutions are best
%% described as programs (i.e. compositions of modular subroutines at
%% various levels of abstraction), 2) most of the solution space is bad
%% in a relatively uninformative way (i.e. it just doesn't work), and
%% thus 3) outcomes are not informative about how you are doing unless
%% you are already quite close to a solution.

%% Intuitively, the way to solve such problems is to find solutions to
%% simple cases in the domain and extract conceptual building blocks that
%% can be composed to scale up to complex cases. With experience, the
%% agent may be able to discover local structure in the solution space,
%% and utilize more traditional AI and machine learning techniques. In
%% this work, we will be focusing on the first phase of this learning
%% problem.

%% %% Of course, every learner

%% %% Choosing the wrong representation, in
%% %% which local structure is not apparent, is often the difference between
%% %% a successful learner and an unsuccessful one. In many respects, this
%% %% may be the kind of situation that adult humans face when confronting
%% %% new domains: we have a lot of general knowledge of successful
%% %% representations, useful strategies, etc.

%% %% But suppose that you have very little knowledge of a domain. Your
%% %% intial representation of the domain is poor: ``nearby'' points in the
%% %% solutions space ellicit wildly different outcomes in unpredictable
%% %% ways. Or perhaps nearby points ellicit the same outcome and its not a
%% %% good one, so you don't know which way to go. Many learning problems,
%% %% especially those whose solutions are best described as programs, have
%% %% these qualities: imagine learning to synthesize new sentences

%% %% teness. We choose the representations our learners use
%% %% and these representations generally Imagine you are trying to learn to
%% %% bake via trial and error. You know what kind of ingredients are used
%% %% in baking (flour, water, butter, sugar, yeast, baking soda, etc.) and
%% %% you know a bit about how to combine these ingredient (mix, whisk,
%% %% etc.) and how to apply heat (oven). But baking is a notoriously
%% %% fragile process: ratios of ingredients, the order of mixing, cooking
%% %% temperatures, resting times, all make a huge difference, and the
%% %% outcome of the baking process is mostly all or none. Either the
%% %% outcome is what you expected or quite unlike it. Even if you do create
%% %% something similar to your intent, it is difficult to assign blame for
%% %% the outcome to any one step or decision along the way.

%% %% Many difficult learning problems have this structure -- learning
%% %% language, developing technologies and scientific theories, learning to
%% %% play instruments or execute other complex motor patterns. These are
%% %% problems in which 1) there is a large set of similar tasks; 2) their
%% %% solutions are best described as programs; 3) the solution space is
%% %% not smooth with respect to the outcome; 4) and the outcome is generally
%% %% uninformative unless you are already pretty close to a solution.

%% %% Of course, knowledge is not acquired wholly by trial and error. But in
%% %% many domains, there is at least initially such a period of ignorance,
%% %% and the question we are trying to tackle here is how you ever get off
%% %% the ground.

%% I need help here.

%% %% Let a \emph{task} be a function $t : \mathcal{L} \rightarrow \{0,
%% %% 1\}$, where $\mathcal{L}$ is the set of all expressions in our
%% %% representation language. We say that $e \in \mathcal{L}$ \emph{solves}
%% %% task $t$ if $t(e) = 1$. Our goal is to solve as many tasks as
%% %% possible, as efficiently as possible. That is, suppose that there is a
%% %% distribution over tasks $t \sim  p_t(\cdot)$, we would like an ordering $O
%% %% = e_1, \dots $ on expressions in $\mathcal{L}$ such that if $p_t(t_1)
%% %% > p_t(t_2)$ then the first expression in $O$ to solve $t_1$ comes
%% %% before the first expression in $O$ to solve $t_2$.

%% %% We motivate a solution to this problem by considering a heirarchical
%% %% Bayesian formulation of the task distribution. Suppose we believe that
%% %% there is some underlying structure to $p_t(\cdot)$ that is explained by
%% %% the fact that for each task $t$ there is a latent expressions $e \in
%% %% \mathcal{L}$ drawn from a stochastic grammar $\mathcal{G}$ and that

%% %% \begin{align}
%% %% e &\sim p_\mathcal{L}(\cdot) = \mathcal{G}\\
%% %% t | e &\sim \text{U}\{ \text{tasks solved by }e\}.
%% %% \end{align}

%% %% Then $p_t(t) = \sum_{e_j \in \mathcal{L} \text{ s.t. } t(e_j) = 1}
%% %% p_\mathcal{L}(e_j)$. If we make the assumption that for the set $E$ of
%% %% expressions in this sum, there is some expression $e^*$ such that
%% %% $p_\mathcal{L}(e^*) >> p_\mathcal{L}(e), e \neq e^*$, then we have, approximately, 
%% %% \[ p_t(t) \propto} p_\mathcal{L}(e^*). \]

%% %% This suggests a way to find the ordering $O$: infer a stochastic
%% %% grammar $G$, under the assumption that $G$ favors one solution to a
%% %% given task over any other, and order the elements of $\mathcal{L}$
%% %% according to $G$.

%% %% Thus, our basic procedure will be to search for a single ``good''
%% %% solution for as many tasks as we can, and to use those solutions to
%% %% estimate a stochastic grammar. Our search will be performed in the
%% %% order of the most recently inferred grammar. So as our ordering
%% %% becomes more efficient and we can solve more tasks, we learn better
%% %% stochastic grammars.

%% %% Formally then, our goal is to solve the following minimal feedback
%% %% multi-task program induction problem: suppose we are given set of
%% %% \emph{tasks} $T=\{t_k\}_{k=1}^K$ where each task is a function $t_i :
%% %% \mathcal{L} \rightarrow \{0, 1\}$, where $\mathcal{L}$ is the set of
%% %% programs in our language.  We say that $ p \in \mathcal{L}$
%% %% \emph{solves} $t_i$ if $ t_i(p) = 1$. Our goal is to find a set of
%% %% solutions $X = \{x_k\}_{k=1}^k \subset \mathcal{L}$ such that

%% %% \[
%% %% X = \argmax_{x_1, \dots, x_K \in L} \sum_k t_k(x_k).
%% %% \]

%% Based on the motivation set out above, we present an iterative
%% strategy. Given a stochastic grammar $\mathcal{G}$ which defines a
%% distribution over the primitives from $\mathcal{L}$ and a integer
%% \emph{frontier size} $N$, we perform two steps in each iteration.  In
%% the first step, we explore the $N$ highest probability programs
%% according to $\mathcal{G}$. In the second step, we compress the
%% programs that solve any of our tasks, and we use that compression to
%% define a new grammar. Repeating these two steps, we effectively
%% reparameterize our search space to include programs that are more likely
%% to be solutions to our tasks.

The remainder of this paper will include two major sections.  We will
first describe the representation language we chose and the details of
the exploration and compression steps of the algorithm. We will then
present results from two experiments involving the domains of
symbolic regression and boolean circuit learning.

%% \section{}
%% Let $\mathcal{F}$ be a family of distributions over $\mathcal{L}$ such
%% that for any $\mathcal{D} \in \mathcal{F}$ we can a) evaluate the
%% probability of program $p$, $P_D(p)$, and we can b) enumerate elements
%% from $D$ in descending order.

%%  To motivate a solution to our problem, we begin with the hierarchical
%% Bayesian formulation of multi-task program induction found in Liang et
%% al.~\cite{liang10programs}. The general idea there is to estimate
%% solution programs for a set of related problems by inferring the posterior
%% distribution over solutions under the assumption that all the solution
%% programs are generated from a common latent stochastic grammar. By
%% jointly inferring solutions and the latent grammar, we find a set of
%% solutions that favors reuse of subcomponents.

%% While this framework is appealing in its elegance and Bayesian
%% interpretation, it cannot be straightforwardly adapted to the
%% situation we are considering here because their proposed MCMC
%% algorithm relies on a local log-likelihood signal, the very kind that
%% we are assuming we do not have access to.

%% To begin, we provide a high level overview of the algorithm and the
%% problem that it solves. The problem setup requires a representation
%% language for programs and a family of distributions over expressions
%% in that language, but we leave these details for the following section

\section{Problem statement and terminology}
Our goal is to solve the following minimal feedback multi-task program
induction problem: suppose we are given set of \emph{tasks}
$T=\{t_k\}_{k=1}^K$ where each task is a function $t_i : \mathcal{L}
\rightarrow \{0, 1\}$ and where $\mathcal{L}$ is the set of programs
in our language.  We say that $ p \in \mathcal{L}$ \emph{solves} $t_i$
if $ t_i(p) = 1$. Our goal is to solve as many of the tasks in $T$ as
we can.

We will use the following terminology. A \emph{frontier} will be the
finite set of programs that the algorithm is considering at any one
time. We will say that a task is \emph{hit} by the frontier if there
is a program in the frontier that solves it.

\section{Extracting modular concepts from functional programs}

In order to solve our stated problem, we need to choose a solution
representation language and a compression criterion, and we need to
describe the frontier enumeration and compression procedures.

\subsection{Representing typed functional programs as binary trees}

Following~\cite{liang10programs} and~\cite{Briggs:2008}, we use a
simply typed combinatory logic -- a variable-free subset of the
polymorphic simply-typed lambda calculus~\cite{Pierce_2002} -- as our
program representation language. In short, the combinatory logic
introduces a \emph{basis} of several primitive \emph{combinators} such
that any function in the lambda calculus can be alternatively written
as applications of the basis combinators. Some common basis
combinators are defined as follows:
\begin{align}
I\; x &\rightarrow  x \\
S\; f\; g\; x\; &\rightarrow (f\; x)\; (g\; x) \\ 
B\; f\; g\; x\; &\rightarrow (f\; x)\; g \\ 
C\; f\; g\; x\; &\rightarrow f\; (g\; x) \\ 
\end{align}

The basis combinators are theoretically enough to express any Turing
computable function, but we will assume that our language has a
variety of primitives and, together with the basis combinators, we
call these the primitive combinators. Note that by default we consider
all primitives to be in their ``curried'' form (e.g. a function like
``+'' adds two numbers $x$ and $y$ by being first applied to $x$,
returning the function $(+\;x)$ and then being applied to $y$,
returning the object $(+\;x)\;y$).

The basis combinators can themselves been expressed in the lambda
calculus (with definitions following directly from the equations
above). The lambda calculus has two basic operations -- application
and abstraction -- but in using the combinatory logic we sequester
uses of the abstraction operation inside the combinators. In doing so,
we have replaced the variable binding of the $\lambda$ operator with
the variable \emph{routing} of these basis combinators. Our
representation thus becomes variable-free. See Liang et
al.~\cite{liang10programs} for a more detailed discussion of this
routing interpretation.

Using combinatory logic is very convenient for program synthesis:
since every expression is the application of one expression to another
-- with this recursion bottoming out at the primitive combinators --
each program is a binary tree. Most importantly, any subtree is itself
a well-formed expression; this is not the case in the lambda calculus,
since abstraction introduces long range dependencies between the
$\lambda$ operator and the variables to which that operator refers. In
the lambda calculus, then, a subtree might have free variables, not
bound to any enclosing $\lambda$.

As a simple example of our representation, consider the squaring
function $\lambda x. * x\; x $. Using two of the basis combinators
above, we can write the squaring function in combinatory logic as $S *
I$ (taking $*$ as a primitive), where application associates to the
left, so we drop the parentheses. When we apply this combinator to a
value $x$, the action of the combinator is defined as $S \;*\; I\;
x~\rightarrow~(*\; x)\; (I\; x)~\rightarrow(*\; x)\; x$.

We can extend this representation with a simple polymorphic type
system~\cite{Pierce_2002}. In this representation, a type is either a
type primitive (e.g. reals, integers, booleans, etc.), a type variable
(e.g. $\sigma$), or a function type $ t_1 \rightarrow t_2$ of
functions from source type $t_1$ to target type $t_2$. Any combinator
can be represented as a binary tree~(Figure~\ref{fig:clbintree}) whose
leaves are typed primitive combinators and whose interior nodes
represent typed applications of combinators (we annotate interior
nodes with their types).

\begin{figure}
%%  \centering
  \begin{subfigure}[]{0.45\linewidth}
    \begin{tikzpicture}[scale=.8,
        transform shape, level/.style={sibling distance=5mm/#1, font=\scriptsize}
        ]
      \node [font=\scriptsize] (top) {$\mathbb{R}$}
      child {node [above=.3cm, left=.1cm, text width=1.8cm] (a) {$(\mathbb{R} \rightarrow \mathbb{R}) \rightarrow \mathbb{R} 
          \rightarrow \mathbb{R}$}
        child {node [left=.5cm, below=1cm, text width=3.0cm] 
          {$ \mathbf{S} : (t_2 \rightarrow t_1 \rightarrow t_0)  \rightarrow (t_2 \rightarrow t_1) 
            \rightarrow \\ t_2 \rightarrow t_0$}}
        child {node [right=0cm, text width=2cm] {$ \mathbf{*}: \mathbb{R} \rightarrow \mathbb{R} \rightarrow \mathbb{R}$}}
      }
      child {node [below right=-.5cm and .2cm] (b) {$\mathbb{R} \rightarrow \mathbb{R}$}
        child {node [above=.5cm] {$ \mathbf{I}: t_0 \rightarrow t_0 \rightarrow t_0$}}
      }
    \end{tikzpicture}
    \caption{}
    \label{fig:clbintree}
  \end{subfigure}
  \begin{subfigure}[R]{0.5\linewidth}
    \begin{tikzpicture}[scale=.7, transform shape, 
        level/.style={sibling distance=10mm/#1},
        every edge/.style={draw, thin}
      ]
      \node [circle,draw] (a) {}
      child {node [left=0cm] (b) {$f_1$}}
      child {node [left=.2cm] (c) {\dots} edge from parent[draw=none] }
      child {node [left=.4cm] (e) {$f_n$}}
      child {node [circle,draw, left=1cm] (d) {} edge from parent[very thick]
        child {node [circle, draw, left=.25cm] (l) {} edge from parent[draw=none]
          child {node [left=.5cm] (b) {$f_1$}} 
          child {node [left=.3cm] (c) {\dots} edge from parent[draw=none] }
          child {node [left=.3cm] (e) {$f_n$} edge from parent[very thick]}
          child {node [circle, draw, left=.25cm] (and2) {}
            child {node [circle, draw, left=.25cm](l2) {} edge from parent[draw=none] 
              child {node [above=.5cm] {\vdots} edge from parent[draw=none]}
            }
            child {node [circle, draw] (r2)  {} edge from parent[draw=none] 
              child {node [above=.5cm] {\vdots} edge from parent[draw=none]}
            }
          }
        }
        child {node [circle, draw, right=.25cm] (r) {} edge from parent[draw=none]
          child {node [circle, draw, right=.25cm] (and3) {}
            child {node [circle, draw ](l3) {} edge from parent[draw=none] 
              child {node [above=.5cm] {\vdots} edge from parent[draw=none]}
            }
            child {node [circle, draw, right=.2cm ] (r3)  {} edge from parent[draw=none] 
              child {node [above=.5cm] {\vdots} edge from parent[draw=none]}
            }
          }
          child {node [right=.3cm] (b) {$f_1$} edge from parent[very thick]} 
          child {node [right=.4cm] (c) {\dots} edge from parent[draw=none] }
          child {node [right=.6cm] (e) {$f_n$}}
        }
      };

      \path (d)  edge[very thick]  node[inner sep=0mm,pos=.4] (a1) {} node {} (l);
      \path (d)  edge[very thick]  node[inner sep=0mm,pos=.4] (b1) {} node {} (r);
      \path (a1) edge[bend right] (b1) ;

      \path (and2)  edge node[inner sep=0mm,pos=.4] (t0) {} node {} (l2);
      \path (and2)  edge node[inner sep=0mm,pos=.4] (k0) {} node {} (r2);
      \path (t0) edge[bend right] (k0) ;

      \path (and3)  edge node[inner sep=0mm,pos=.4] (t1) {} node {} (l3);
      \path (and3)  edge node[inner sep=0mm,pos=.4] (k1) {} node {} (r3);
      \path (t1) edge[bend right] (k1) ;
    \end{tikzpicture}
    \caption{AND/OR expression search space. }
    \label{fig:andor}
    \end{subfigure}
  \caption{(a) The typed combinator $S * I$ ($f(x)=x^2$) represented as
    binary tree. (b) The space of all typed combinators represented as an
    infinitely deep AND/OR tree. Any specific combinator is a
    sub-binary tree of this AND/OR tree whose leaves consist of only
    the $f_i$ nodes (i.e., the primitive combinators of the grammar).}

\end{figure}

\subsection{A stochastic grammar over programs.}
\label{sec:stochgrammar}
We would like a way to describe how likely we believe a program to be
given the probabilities we ascribe to its components. Recall that
$\mathcal{L}$ is the set of all programs.  A stochastic grammar over
programs represents a distribution over expressions in
$\mathcal{L}$. We associate with each primitive combinator $c_i$ in
$\mathcal{L}$ a probability $p_i$, $0~\leq~p_i~\leq~1, \sum_i~p_i~=~1
$.

The probability of using a primitive combinator somewhere in a program
should depend on which other combinators could have been used in its
place. This, in turn, depends on the type at that particular location
in the program tree; we call this type the \emph{requesting} type. For
example, suppose that we have the primitive $(+1)$ of type $Int
\rightarrow Int$ and the primitive $NOT$ of type $Bool \rightarrow
Bool$. If the requesting type is $\sigma \rightarrow Int$, then we can
only use $(+1)$, but if the requesting type is $\sigma \rightarrow
\sigma$ then we could use either $(+1)$ or $NOT$. Let $C = c_1, \dots,
c_n$ be the set of primitive combinators in our grammar; we define
$C_t$ for any type $t$ to be the set of combinators in $C$ that are
consistent with $t$, that is, whose types \emph{unify} with t (we use
unification according to the standard type theory
presentation~\ref{Pierce_2002}).

We define the probability of an expression $e$ from a grammar with
primitive $C$ to be $p(e)~=~\prod_{n=1}^N p(c_n | t(c_n))$ where
$p(c_n | t(c_n))$ is the probability of using primitive $c_n$ when the
requesting type was $t(c_n)$. This definition means that the
probability of an expression only depends on the probabilities
assigned to the library primitives. We define these conditional
probabilities as
\[
p(c_i | t(c_n)) =    \frac{1}{\epsilon}\frac{p_i}{\sum_{C_{t(c_n)}} p_j}.
\]
That is, $p(c_i | t(c_n)$ is the probability of sampling $c_i$ from
the multinomial distribution defined by the $p_i$'s, conditioned on
the requesting type, with the factor of $\frac{1}{\epsilon}$ added as
a simple way to ensure that the sum over the probabilities of all
expressions converges to a finite value (so that the distribution over
expressions is proper). This is true whenever $\epsilon \geq
4$\footnote{The probability mass $M_d$ of all expressions of depth
  less than or equal to $d$ can be written with the recurrence
  relation $M_d \leq M_{d-1}^2 + M_1$. This has an asymptotic fixed
  point, as $d \rightarrow \infty$ if $x = x^2 + M_1$ has a solution,
  which is true if $M_1 \leq 1/4$.}.

To estimate the values $p_i$ from a set of expressions, we simply divide
the number of times primitive combinator $c_i$ was used by the number
of times it could have been used given the requesting types in the
expression trees. Note that this is not a maximum likelihood estimator
for this distribution; finding that would require solving a rather
complex optimization problem because of each choice is conditioned on
the requesting type. We choose our estimator merely for convenience,
and find that our empirical results justify this simplicity. It is a
matter for future research to determine to what extent a more accurate
estimator would improve performance.

\subsection{Best-first enumeration of most probable programs.}
In order to explore the frontier of most promising programs, we need a
procedure that enumerates the $N$ most probable programs. There has
been recent interest in this problem, most of it focused on
enumerating the shortest program satisfying some
criterion~\cite{DBLP:conf/sfp/Katayama05}
\cite{DBLP:conf/aaip/YakushevJ09}. Our approach is to formulate the
problem as a best-first exploration of an AND/OR tree
(Figure~\ref{fig:andor})~\cite{nilsson1982principles}~\cite{DBLP:journals/cacm/Hall73},
which we denote $\mathcal{G}$. In words, the enumeration procedure is
best described by the following recursion: every program is either a
primitive combinator or it a left child program applied to a right
child program.

This translates to the following AND/OR tree. Suppose we want a
program of type $t$. The root of $\mathcal{G}$ is an OR node of type
$t$. Its children are elements in $C_t$ (defined in the previous
section) and one AND node of type $t$ with two children. Each child of
an AND node has the same structure as the root node, with modified
types: type of its left child is $\tau \rightarrow t$ (where $\tau$ is
a fresh type variable). Since the type of the right child is
constrained by the subtree rooted at the left child, we always expand
the left child first. Once we have a complete left child program, we
can use its target type as the type of the right child.

Recall that a valid partial solution $\mathcal{H}$ is a subtree of
$\mathcal{G}$ satisfying the following properties: the root of
$\mathcal{H}$ is the root $\mathcal{G}$ and any OR node in
$\mathcal{H}$ has at most one child. $\mathcal{H}$ is a complete
solution (and thus a complete program) if it is a partial solution
whose leaves are leaves of $\mathcal{G}$, if each OR node in
$\mathcal{H}$ has exactly one child, and if each AND node in
$\mathcal{H}$ is connected to all of that node's children in
$\mathcal{G}$

The value of a partial solution $\mathcal{H}$ is calculated by summing
the value of its leaves, where each such value is the log of the
probability assigned to that leaf in
Section~\ref{sec:stochgrammar}. If a leaf of $\mathcal{H}$ is a leaf
of $\mathcal{G}$ then either it is an AND node, to which we assign a
value of 0, or it corresponds to some primitive combinator $c_i$ in
the library and is the child of some typed OR node with type $t$. We
assign it a value of $\log{p(c_i | t)}$. If a leaf $v$ of
$\mathcal{H}$ is an OR node, then the value of any extension of it to
a complete solution is at most $log(1/4)$. Thus, we have defined a
value for partial solutions that upper bounds the value of any of its
extensions.

Our best-first policy is to expand the next partial solution
with the highest value. If a partial solution has multiple leaves that
are not primitive combinators, we expand the left most one first,
since it is the choice of left combinator that defines the type of its
right combinator in an application pair. That is, the parent node
applies its constraint on valid solutions via the left child, not the
right child. We want to apply constraints as early as possible, so
expanding left children first prunes more partial solutions.

\subsection{Finding the most compressive set of solutions}

Having enumerated the frontier, we want to assign solutions
to hit tasks such that this set is maximally compressible. This will
promote the usage of reusable modular subtrees.

A natural metric for the compressibility of a set of binary trees is
the number of unique subtrees (counting leaves as subtrees) that this
set contains. As an example of this, suppose we had six tasks whose
solutions were integers $4^2, 5^2, 6^2, 7^2, 8^2$ and suppose our
frontier contained the 10 expressions:

\captionsetup[subfigure]{labelformat=empty} 
\begin{figure}[h!]
   \begin{subfigure}{.17\linewidth}
     \begin{tikzpicture}[scale=.7, background rectangle/.style={fill=lightgray}] 
       \tikzset{level distance=15pt, sibling distance=10pt}
       \Tree [[  * 4 ] 4 ]
     \end{tikzpicture}
  \end{subfigure}
  \begin{subfigure}{.17\linewidth}
     \begin{tikzpicture}[scale=.7]
       \tikzset{level distance=15pt, sibling distance=10pt}
       \Tree [[  * 5 ] 5 ]
     \end{tikzpicture}
  \end{subfigure}
   \begin{subfigure}{.17\linewidth}
     \begin{tikzpicture} [scale=.7]
       \tikzset{level distance=15pt, sibling distance=10pt}
       \Tree [[  * 6 ] 6 ]
     \end{tikzpicture}
  \end{subfigure}
   \begin{subfigure}{.17\linewidth}
     \begin{tikzpicture}[scale=.7] 
       \tikzset{level distance=15pt, sibling distance=10pt}
       \Tree [[  * 7 ] 7 ]
     \end{tikzpicture}
  \end{subfigure}
  \begin{subfigure}{.17\linewidth}
     \begin{tikzpicture}[scale=.7] 
       \tikzset{level distance=15pt, sibling distance=10pt}
       \Tree [[  * 8 ] 8 ]
     \end{tikzpicture}
  \end{subfigure}
  \vspace{2mm}

  \hdashrule{.9\linewidth}{1pt}{1pt}

  \vspace{2mm}
  \begin{subfigure}{.17\linewidth}
     \begin{tikzpicture}[scale=.6] 
       \tikzset{level distance=15pt, sibling distance=10pt}
       \Tree [[[  S * ] I ] 4 ]
     \end{tikzpicture}
     \caption{(16)}
  \end{subfigure}
  \begin{subfigure}{.17\linewidth}
     \begin{tikzpicture}[scale=.6]
       \tikzset{level distance=15pt, sibling distance=10pt}
       \Tree [[[  S * ] I ] 5 ]
     \end{tikzpicture}
     \caption{(25)}
  \end{subfigure}
   \begin{subfigure}{.17\linewidth}
     \begin{tikzpicture} [scale=.6]
       \tikzset{level distance=15pt, sibling distance=10pt}
       \Tree [[[  S * ] I ] 6 ]
     \end{tikzpicture}
     \caption{(36)}
  \end{subfigure}
  \begin{subfigure}{.17\linewidth}
     \begin{tikzpicture}[scale=.6] 
       \tikzset{level distance=15pt, sibling distance=10pt}
       \Tree [[[  S * ] I ] 7 ]
     \end{tikzpicture}
     \caption{(49)}
  \end{subfigure}
  \begin{subfigure}{.17\linewidth}
     \begin{tikzpicture}[scale=.6] 
       \tikzset{level distance=15pt, sibling distance=10pt}
       \Tree [[[  S * ] I ] 8 ]
     \end{tikzpicture}
     \caption{(64)}
  \end{subfigure}
\end{figure}
\captionsetup[subfigure]{labelformat=parens}

The top row consists of one type of representation of an integer
squared, namely, the integer times itself. The bottom row contains
another type of representation in which the function $f(x)=x^2$,
represented in combinatory logic as $S * I$, is applied to the
integer. The top and bottom row of each expression evaluates to the
same integer. If we consider only the first column, the top tree has 4
unique subtrees and the bottom has 7. But each column adds 3 unique
subtrees to the top row and only 2 to the bottom. So by the fifth
column the top row has more unique subtrees than the bottom row. The
increased compressibility of the bottom row is due to the fact that
the squaring operation has been factored out as an abstraction. This
is an example, then, of how this compressibility metric favors
abstraction when multiple expressions are being considered at once.

We want to choose a solution for each hit task such that the complete set
of such solutions for all solved tasks is as compressible as
possible. That is, let $t_i$ be a task in the set of hit tasks, and
let $\{e^k_i\}$ be the expressions in the frontier that solves
it. Then our goal is to find
\[
\{e^*_i\} = \argmin_{\{e^k_i\}} | \cup_i e^k_i |,
\]
where $|\cdot|$ denotes the number of unique trees in a set of expressions.

However, an exact optimization of this problem requires examining
$O(M^K)$ assignments, where $M$ is the maximum number of solutions for
any one task and $K$ is the number of tasks. Since this is
prohibitive, we relax the compressibility metric as follows: we define
the cost of adding each new solution $s_n$ to a partial set of
solutions $s_1, \dots, s_{n-1}$ to be the number of additional unique
subtrees in $s_n$ that are not present in $s_{n-1}$. That is, the
penalty of adding a new expression to the set of expressions depends
only on the last expression we added. Our solution thus becomes
approximate and that approximation is order-dependent, but a
depth-first search on the defined search space goes from exponential
in the solution degree to quadratic ($O(KM^2)$).

However, when the number of solutions in the frontier grows, this
quadratic computation becomes prohibitive. Therefore, in practice, we
bound the number of potential expressions we are willing to consider
at any one iteration to any one task to a small number. We find that a
setting of this bound to 2 is computationally feasible and large
enough to guide subcomponent discovery in a reliable way. It may be
necessary to increase this bound as we scale this algorithm to larger
problems.

\subsection{Re-estimating the stochastic grammar over programs.}

Given a set of solution programs, we want to re-estimate our
stochastic grammar. Our inspiration for doing this is based on
the Neville-Manning algorithm for grammar-based
compression~\cite{nevill1997identifying}. The idea of that compression
algorithm is to compress any string by creating a grammar for that
string such that a) no two symbols appear more than once in the
grammar and b) every rule is used more than once. From a set of trees
we generate a grammar according to these criteria.

%% To do this, we traverse a set of combinator binary trees in
%% depth first order. We count the occurrences of every subtree. Every
%% time we encounter a subtree for the \emph{second} time, we designate
%% it as a primitive in our new grammar. This ensures that every
%% primitive is used more than once. During our traversal, we do not
%% descend into trees that have been added already to the new
%% grammar. Thus, we only count an instance of a tree if that instance is
%% not a subtree of a primitive element of the new grammar.

This procedure generates a parse of the solution set with a new set of
primitives, and we keep count of how many times each primitive
occurred in the solutions set. To estimate the probabilities
associated with each node production, we again traverse the solution
set, but this time for each node we count which other primitive
elements from the new grammar could have been used. In accordance with
Section~\ref{stochgrammar}, we estimate the terminal production weight
to be the ratio of the number of times that node was used in the first
traversal divided by the number of times it was used in the second
traversal.

\section{Experiments.}
\subsection{Symbolic regression.}
\label{sec:symreg}

In this section, we discuss the performance of our learning algorithm
on a symbolic regression problem. In this problem, each task is a set
of pairs $ (x, f(x))$ where $x \in (0, \dots, 9)$ and $f$ is a
polynomial. The goal is to discover a program for each task that
outputs the appropriate value $f(x)$ for each input x.

%% The choice of an initial grammar for the algorithm will, of course,
%% have a big effect on its performance, and so we would like to
%% characterize this effect. We can divide primitives into roughly two
%% sets; there are low-level primitives, those that operate on concrete
%% types like Bool or Double, and these tend to be domain specific. For
%% example, concrete primitives include $1, 2$, etc., and $f(x) = 3x$. The
%% other class of primitives are abstract primitives, which can be
%% characterized by having arguments whose types are primarily function
%% types, like $(t \rightarrow s), (t \rightarrow s) \rightarrow \text{
%%   Double },$ etc. Of course, we can construct operators that fall in
%% between.

%% The most abstract operators are those that Liang et. al refer to as
%% ``routers''~\cite{liang10programs}. These operators take the place of
%% $\lambda$ abstraction in the general lambda calculus: whereas
%% $\lambda$ abstraction allows us to bind arguments to variables nested
%% deep inside an expression, without them we need to have operators that
%% ``route'' arguments to the appropriate functions in a nested
%% expression.

%% \begin{definition}[Router~\cite{liang10programs}] A \emph{router} is a combinator
%% represented by a finite sequence of elements from $\{B, C, S\}$. The
%% behavior of router $r$ is given by
%% \[
%% (r \, f \, g \, x_1 \dots x_{|r|}) = (f \, x_{i_1} \dots x_{i_n}) (g
%% \, x_{j_1} \dots x_{j_n}),
%% \]
%% where $i_1 < \dots < i_n$ are indices $i$ such that $r_i \in \{C, S\}$
%% and where $j_1 < \dots < j_n$ are indices $j$ such that $r_j \in \{B,
%% S\}$. Let $\mathcal{R}_k$ be the set of all routers of length equal to
%% $k$. Let $\mathcal{R}_{\leq k}$ be the set of all routers of length
%% less than or equal to $k$.
%% \end{definition}

%% To analyze the behavior of our algorithm, we will vary the parameters
%% of the problem along two axes while keeping the learning tasks the
%% same. Along the first axis, we vary the initial library to which our
%% learner has access with respect to the its abstractness (i.e., number
%% of routers) and its concreteness (i.e. number of concrete
%% primitives). Along the second axis we vary the size of the search
%% frontier.

Each task is to predict the value of an unknown polynomial on the
inputs $0, \dots, 9$. The set of tasks corresponds to the set of
polynomials $\{ax^2 + bx + c | a, b, c \in \{0, \dots, 9 \}\}$. In the
initial library, we include the basic combinators I, S, B, C and four
arithmetic primitives, $1$, $0$, $*$, and $+$. The choice of initial
weights on these primitives can make a difference, as it determines
the relative ordering of combinators in the enumeration step. To get a
general sense of the algorithm's performance, we set the initial
weights to be slightly different on each run, perturbing them around a
uniform weighting.

\begin{figure}[h]
\centering
\begin{subfigure}[Before]{\linewidth}
\includegraphics[width=.9\linewidth]{figures/symreg3frontiers.pdf}
\label{fig:symregLearning}
\caption{Learning curves as a function of frontier size.  As frontier
  size is increased, it plateaus closer to $100\%$ performance. A
  control brute-force search over $150000$ only hits $3\%$ of the
  tasks.}
\end{subfigure}

\begin{subfigure}[After]{.65\linewidth}
\includegraphics[width=.9\linewidth]{figures/symRegDiffCurricula.pdf}
\caption{How do different task sets affect learning curves? Learning
  curves at frontier size of $10000$ for different task sets. Note
  that performance is equally good if either the constant functions of
  the linear functions are missing, by drops dramatically when both
  are absent, even though these only account for about $10\%$ of the
  tasks. Performance drops even lower if we only consider ``complex'' quadratics.}
\label{fig:symregCurricula}
\end{subfigure}
\begin{subfigure}[After]{.3\linewidth}\
\includegraphics[width=.9\linewidth]{figures/symregCircuitDepths.pdf}
\caption{Circuit depth as a function of iteration on a typical run.}
\label{fig:symregDepths}
\end{subfigure}

\caption{Symbolic regression experiment data.}
\label{fig:symreg} 
\end{figure}

%% Why do the learning curves asymptote early? One might object that a
%% learning algorithm should continuously improve, if only slowly, as it
%% is given more time. But in these runs we observe a finite and
%% permanent performance cap. This is expected: once the learner find a
%% set of concepts it can explain, if those concepts fill up the
%% frontier, then there is nothing to spur the learner to incorporate as
%% yet unsolved tasks.

%% To investigate whether this is in fact the case, we can examine the
%% proportion of expressions in the frontier that match tasks in our task
%% set; that is, is the learning curve plateauing because the frontier is
%% heavily redundant, or is it plateuing because the frontier contains
%% many expressions that do not hit the task set at
%% all? Figure~\ref{fig:plateau_investigation} shows the proportion of
%% the expressions in the frontier that hit a task as the frontier size increases.

In Figure~\ref{fig:symreg}, we show performance results as we vary the
frontier size. Each individual plot shows how changing the frontier
size changes the number of tasks the algorithm identifies correctly
over the course of fifteen algorithm iterations. As a control, we
enumerated $10000*15=150000$ expressions from the initial
grammar. This is the total number of expressions that a run of the
algorithm sees if it has a frontier size of $10000$ and runs for $15$
iterations. Thus, if the benefit accruing to the runs with larger
frontier sizes is due to simple an increase in brute-force search
ability, rather than increased learning, we should see similarly good
performance from this control run. In fact, however, the control run
only hit $27$ of the $1000$ asks ($3\%$), whereas our algorithm nears $100\%$.

We would like to assess to what degree ``conceptual bootstrapping'' is
responsible for the improved performance. In the next experiment we
will inspect directly the modular components that are learned by this
algorithm. Here we look at the idea of a ``learnable'' curriculum,
that is, the notion that conceptual bootstrapping requires that a
set of tasks contain a subset of tasks that are sufficiently simple so
that the learner is able to solve those with its initially
impoverished library of concepts.

In Figure~\ref{fig:symregCurricula}, we present learning curves
(frontier size 10000) corresponding to various versions of our
original symbolic regression task set. Recall that the original set
consisted of all polynomials of degree two or less and with
coefficients between 0 and 9. In the ``no constant'' task set, we
remove the 9 constant functions in this set. In the ``no linear'' task
set, we remove 90 linear functions from the set. We observe that
performance on those task sets does not decrease. However, when we
remove both the linear functions and the constant function (``only
quadratics''), we see a sharp drop in the algorithm's
performance. When we further restrict the task set to only ``complex''
quadratics (which we are defining as quadratics all of whose
coefficients are greater than zero), we observe another comparable
drop in performance. When we go one step further and restrict the task
set to quadratics with coefficients greater than 1, performance drops
to 0.

%% In addition, we have a
%% baseline for each curve, visits the same number of expressions as the
%% corresponding algorithm run (i.e. frontier size $\times$ $\#$ of
%% iterations) but all at once; it does not do any learning. Thus, this
%% baseline serves as a brute-force control for the learning aspect of
%% the algorithm.

\subsection{Boolean circuits}
\label{sec:boolcircuits}

As another demonstration of this approach, we investigate the
procedure's ability to learn boolean functions using only the logical
primitive NAND and the basic combinators. It is well known that a
boolean circuit can be constructed for any boolean function given only
the NAND gate. Here, the basic combinators take the place of the
wiring normally found in a circuit, the output of element to multiple
inputs downstream.

\begin{table}
\tiny
\begin{tabular}[width=\linewidth]{|p{1cm}|p{2cm}|p{3.5cm}|}
\hline \\
  function  & CL expression & circuit  ~ \\ \hline
¬†¬†¬†¬†¬†¬†¬†¬†TRUE¬† & (S NAND) (S NAND I)¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬† & ~

%\begin{top}
\begin{circuitikz}[transform shape] 
\draw
(1, 0) node[nand port, scale=.5] (nand1) {}
(2,.3) node[nand port, scale=.5] (nand2) {}
(0, 0) node[anchor=east]  (x) {x}
(x) -- (.3, 0) -| (nand1.in 1)
(x) -- (.3, 0) -|  (nand1.in 2)
(x) |- (nand2.in 1)
(nand1.out) |- (nand2.in 2)
;
\end{circuitikz}
%\end{top}
\\ \hline
¬†¬†¬†¬†¬†¬†¬†¬†FALSE & (S B (S NAND)) (S NAND I)¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬† &  ~

%% \begin{circuitikz}[transform shape]

%% ;
%% \end{circuitikz}

\\ \hline
¬†¬†¬†¬†¬†¬†¬†¬†NOT¬†¬† & S NAND I¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬† &  ~

\begin{circuitikz}[transform shape]
\draw
(1.2,0) node[nand port, scale=.5] (nand1) {}
(0, 0) node (x) {x}
(x) -| (nand1.in 1)
(x) -| (nand1.in 2)
;
\end{circuitikz}

\\ \hline
¬†¬†¬†¬†¬†¬†¬†¬†AND¬†¬† & ((C B NAND) (B (S NAND I)))¬†¬†¬†¬†¬†¬†¬†¬†¬†¬† & ~

\begin{circuitikz}
\draw
(1.0,0) node[nand port, scale=.5] (nand1) {}
(2.0,0) node[nand port, scale=.5] (nand2) {}
% (0, 0) node (x) {x}
% (0, 0) node (y) {y}
node[above left=.2cm and .7cm  of nand1] (x) {x} %-- (nand1.in 1)
node[below left=.2cm and .7cm  of nand1] (y) {y} %-- (nand1.in 2)
(x) |- (nand1.in 1)
(y) |- (nand1.in 2)
(nand1.out) |- (nand2.in 1)
(nand1.out) |- (nand2.in 2)
;
\end{circuitikz}
 \\  \hline
¬†¬†¬†¬†¬†¬†¬†¬†OR¬†¬†¬† & ((B (C (B (S NAND) NAND))) (S NAND I) & ~

\begin{circuitikz}
\draw
(0,0) node (x) {x}
(0,.7) node (y) {y}

node[nand port, scale=.5, right=1cm of x] (nand1) {}
node[nand port, scale=.5, right=1cm of y] (nand2) {}
node[nand port, scale=.5, below right=.3cm and 1cm of nand2] (nand3) {}
(x) -| (nand1.in 1)
(x) -| (nand1.in 2)
(y) -| (nand2.in 1)
(y) -| (nand2.in 2)
(nand1.out) |- (nand3.in 1)
(nand2.out) |- (nand3.in 2)
;
\end{circuitikz}
 \\  \hline
¬†¬†¬†¬†¬†¬†¬†¬†XOR¬†¬† & ((S (B C (((B (C B NAND)) S) (((B (C B NAND)) S) (B C ((B (B NAND)) NAND)))))) I)¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬† & ~

\begin{circuitikz}
\draw
node[nand port, scale=.5] (nand1) {}
node[nand port, scale=.5, above right=.3cm and 1cm of nand1] (nand2) {} 
node[nand port, scale=.5, below right=.3cm and 1cm of nand1] (nand3) {} 
node[nand port, scale=.5, right=2cm of nand1] (nand4) {} 
node[left=1.2cm of nand2.in 1] (x) {x}
node[left=1.2cm of nand3.in 2] (y) {y}
node[right=.2cm of nand4] (out) {}
(x) -| (nand1.in 1)
(x) -| (nand2.in 1)
(y) -| (nand1.in 2)
(y) -| (nand3.in 2)
(nand1.out) -| (nand2.in 2)
(nand1.out) -| (nand3.in 1)
(nand2.out) |- (nand4.in 1)
(nand3.out) |- (nand4.in 2)
(nand4.out) -- (out)
;
\end{circuitikz}

 \\  \hline
¬†¬†¬†¬†¬†¬†¬†¬†F¬†¬†¬†  & S B (S NAND)¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬† & ~
%% \begin{circuitikz}
%% \draw
%% (0, 0) node[left=.25) (x) {x}

%% ;
%% \end{circuitikz}

\\ 
\hline

\end{tabular}
\caption{Expressions in the induced library on sets of tasks
  corresponding to boolean circuits synthesized from AND, OR, and NOT
  gates, with boolean circuits of 1, 2, or 3 inputs.}
\label{table:boolexpr}

\end{table}

%% \begin{table}[t]
%% \scriptsize
%%    \begin{tabular}{|l|l|l|}
%%     function  & CL expressions & circuit  ~ \\ \hline
%%         TRUE  & (S NAND) (S NAND I)                   & ~ \\ \hline
%%         FALSE & (S B (S NAND)) (S NAND I)             & ~ \\ \hline
%%         NOT   & S NAND I                              & \input{figures/NOT.tex} \\ \hline
%%         AND   & ((C B NAND) (B (S NAND I)))           & ~ \\ 
%%         OR    & ((B (C (B (S NAND) NAND))) (S NAND I) & ~ \\ 
%%         XOR   & ~                                     & ~ \\ 
%%         F     & S B (S NAND)                          & \input{figures/F.tex}
%%     \end{tabular}
%% \end{table}

We constructed two task sets. The first of these was constructed
explicitly to contain familiar module structure. In this set of tasks,
we sampled $1000$ boolean circuits using AND, OR, and NOT gates. To
accomplish this, we first sampled either 1, 2, 3 inputs, then between
1 and 5 gates, randomly wiring the inputs of each new gate to the
output of one of the existing gates in the circuit. We continued this
sampling procedure until we had 1000 ``connected'' circuits, i.e.,
circuits all of whose output are wired to an input except for the last
one (the output gate). This resulted in $1000$ tasks consisting of
$82$ unique boolean functions, with a distribution of tasks as shown
in Figure~\ref{fig:booldistr}. In Figure~\ref{fig:boolLearningCurvesCircuits},
We present learning curves for frontier sizes 100, 500 and 10000,
using the same procedure as in Section~\ref{sec:symreg}.

\begin{figure}
\includegraphics[width=\linewidth]{figures/circuitDistrComp.pdf}
\vspace{-10pt}
\caption{Distribution of boolean functions among (a) the set of 1000
  circuits randomly constructed from AND, OR and NOT gates (see text);
  b) the first 1000 expressions from the grammar over programs
  \emph{before} learning; c) the first 1000 expressions from the
  grammar over programs \emph{after} learning. Gray levels indicate
  percent of instances (circuits/expressions) which evaluate to that
  function.\label{fig:booldistr}}
\end{figure}

This problem is particularly suitable for inspecting the constituents
of the induced grammar $\mathcal{G}$. We might hope that our algorithm
recovers the constituent logic gates that were used to build up the
problems. Table~\ref{table:bool} shows a few of the top ranked
expressions in the library. These include the basic logic gates; we
also include one higher-order expression $F$, to stress that the
representation we are using allows one to build concepts that are more
expressive that just sub-circuits.

In a second experiment, we include all $272$ boolean truth tables with
three or fewer inputs as tasks. This is a more difficult problem. Note
that in the first boolean function problem, there are two kinds of
learning that can take place: first, many expressions map to a single
boolean function, so we need to learn primitives that allow us to span
many different functions instead of allows generating very simple
functions over and over again. The second kind of learning involves
the distribution of the functions themselves. In the first boolean
experiment, that structure is apparent in the distribution in
Figure~\ref{fig:boolhist}. In this second experiment, we remove the
second kind of structure. In Figure~\ref{fig:boolLearningCurvesUniform}, we show
10 runs of the algorithm with a frontier size of $2000$: note how
there are several local minima that majority of the runs get stuck in
with performance around $50\%$, but several of the runs seem to take
different trajectories to more successful representations.

\begin{figure}[t]
\begin{subfigure}[Before]{\linewidth}
\includegraphics[width=.9\linewidth]{figures/booleancircuits3frontiers.pdf}
\label{fig:boolLearningCurvesCircuits}
\caption{Learning curves for various frontier sizes on task set of
  Boolean functions drawn randomly from a distribution (see text for
  details) of circuits built up of NOT, AND, OR gates.}
\end{subfigure}

\begin{subfigure}[Before]{\linewidth}
\includegraphics[width=.9\linewidth]{figures/booleanfunctionsUniform.pdf}
\label{fig:boolLearningCurvesUnifrom}
\caption{10 learning curves for frontier size of 2000 on a task set
  consisting of all Boolean functions of cardinality 3 or
  smaller. Note how most of the trajectories get stuck in a local
  minimum around $50\%$.}
\end{subfigure}
\label{fig:boolLearningCurves} 
\caption{Learning curves from Boolean function learning experiment. }
\end{figure}

%% \subsection{Digital arithmetic}
%% Having demonstrated the behavior of the E-C algorithm on symbolic
%% regression, we analyze its performance on tasks involving digital
%% arithmetic, a class of problems with known modular structure. The
%% general goal is to discover boolean circuits consisting of a limited
%% set of gates that perform arithmetic operations on binary number
%% representations.

\section{Discussion}
This work attempts to bring together ideas from various areas of
research -- hierarchical Bayes, learning to learn, minimum description
length learning -- to provide a proof of concept that learning
abstractions from programs via compression can rapidly improve search
performance. We emphasize this by focusing on domains in which the
solution space is not smooth and the error signal is all or
none. Thus, the main source of information for learning is the modular
and compositional structure of the solution space.

Of course, the representations that people use to learn in the world
often have a partially smooth or local structure. Techniques in AI and
machine learning commonly use smoothness or local structure to make
search tractable. But the difference between this kind of learning and
conceptual learning is undeniable. The former is what the statistician
or machine learning researcher implements. The latter is what she
learns as a student and as a practicioner. As researchers, we try
different models, watch them fail or succeed, and abstract out the
elements of these models that capture the relevant differences. We mix
and match parts of different ideas, and utilize succcessful patterns
of reasoning that can span many problems and domains.

We believe that the primary characteristic of conceptual learning is
that it occurs over a modular and compositional computational
representation that supports higher-order functions. That is why we
chose a functional programming language as the representational
language of this work.

The experiments in this work do not contain the richness of real world
learning, but we believe that the learning algorithm's behavior
exhibits some of the characteristics we would expect of a conceptual
learner.

One characteristic of this algorithm's behavior is a rapid initial
increase in performance followed by a hard plateau. This plateau is
expected given the circular nature of the algorithm: the learner
inspects the region of the solution space that is typical of solutions
already found. We see in this phenomenon a conceptual demarcation of
the problem space: a coherent set of concepts or problems solving
techniques can define by their power the extent of problems a problem
solver is willing to consider. One might hope, however, that an
intelligent learner would recognize that such a demarcation has
occured and attempt to find another class of concepts for the
remaining unsolved tasks.

Another aspect of conceptual learning that we see present in this
algorithm's behavior is the importance of a curriculum. By this we do
not mean an explicit ordering of the problems in the task set, but
rather that a set of sufficiently simple problems must act as stepping
stones to the acquisition of complex representations.

%% The file named.bst is a bibliography style file for BibTeX 0.99c
\bibliographystyle{plain}
\bibliography{ijcai13}
%% \nocite{*}

\end{document}