### 2013-06-10 drafts/inference/inference.tex
Added a writeup of how EM and variational inference can be used to perform inference in our generative model

\documentclass[12pt]{article}
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{amsfonts}

\begin{document}
\section{EM Inference}
E Step: Compute $q(\{e_i\})=P(\{e_i\} | \{t_i\},G^{old})\propto P(\{t_i\}|\{e_i\})P(\{e_i\}|G^{old})$.
Observe that $q(\{e_1,e_2,\cdots\})=q(\{e_1\})q(\{e_2\})\cdots$

M Step: Find $G$ by
\begin{eqnarray*}
&\max_G& \mathbb{E}_q \left[ \log P(\{t_i|e_i\}) P(\{e_i\}|G)P(G|\theta) \right]\\
&\max_G& \log P(G|\theta) + \mathbb{E}_q \left[ \log P(\{e_i\}|G)\right] \\ 
&\max_G& \log P(G|\theta) + \sum_i \mathbb{E}_q \left[ \log P(e_i | G) \right]
\end{eqnarray*}
Consider each term in the sum on the r.h.s.:
\begin{eqnarray*}
\mathbb{E}_q \left[ \log P(e_i | G) \right] &=& \sum_{e_1}\cdots\sum_{e_i}\cdots \sum_{e_n} q(\{e_1,\cdots\})\log P(e_i|G)\\
&=& \sum_{\{e_{j\not=i}\}} q(\{e_{j\not= i}\})\sum_{e_i}q(e_i)\log P(e_i|G)\\
&=& \sum_{\{e_{j\not=i}\}} P(\{e_{j\not= i}\}|G^{old}, \{t_{j\not= i}\})\sum_{e_i}q(e_i)\log P(e_i|G)\\
&=& \sum_{e_i}  P(t_i|e_i)P(e_i|G^{old})/P(t_i|G^{old}) \log P(e_i|G)\\
&=& \mathbb{E}_{q_i} \left[ \log P(e_i|G) \right]
\end{eqnarray*}
Back to what we're maxing,
\begin{eqnarray*}
&\max_G& \sum_i \mathbb{E}_{q_i} \left[ \log P(e_i|G) \right]
\end{eqnarray*}
Draw from $q_i$ using importance sampling: draw from $P(\cdot | G^{old})$, and then weight by $P(t_i | \cdot)$.
Let the draws be $e_i^k$ and the corresponding weights be $w_i^k$.
Approximate the expectation by using this sample; now the maximized quantity is
\begin{eqnarray*}
&\max_G& \sum_i \sum_k w_i^k \log P(e_i|G)\\
&\max_G& \prod_{i,k}  \left( P(e_i|G) \right)^{w_i^k}
\end{eqnarray*}
This is a nice form: what we're maximizing is the probability of generating a set of parse trees, each of which is observed in proportion to $w_i^k$, which is proportional to $P(t_i|e_i)$.
We can perform this max using hill climbing or some sort of stochastic search, completing the ``M'' step of EM.

\section{Variational Inference}
It would be nice to represent a distribution over grammars.
This would, for example, allow generalization from a single task, permitting one-shot program induction.
It also would make our inference algorithm closer to the actual generative model.

As a first attempt, I'll approximation the posterior $P(\{e_i\},G | \{t_i\})$ using a factorized distribution (mean-field approximation):
$$
Q(\{e_i\},G) = q(G) \times \prod_i q(e_i) 
$$
where $Q$ is our approximation to the posterior.
Following the standard derivation of variational Bayes, we want to solve the equations
\begin{eqnarray*}
\log q(e_i) &\stackrel{+}{=}& \sum_G q(G)\sum_{e_{j\not= i}} q(\{e_j\}) \log P(e_i, t_i, \{e_{j\not= i}\}, \{t_{j\not= i}\}, G) \\
\log q(G) &\stackrel{+}{=}& \sum_{\{e_i\}} q(\{e_i\}) \log P(\{e_i\}, \{t_i\}, G )
\end{eqnarray*}
The second equation, of course, gives the same distribution over $G$ as EM:
\begin{eqnarray*}
\log q(G) &\stackrel{+}{=}& \sum_{\{e_i\}} q(\{e_i\}) \log P(\{e_i\}, \{t_i\}, G )\\
 &\stackrel{+}{=}& \sum_{\{e_i\}} q(\{e_i\}) \log P(\{e_i\} | G )P(G|\theta)\\
 &=& \log P(G|\theta) + \sum_{e_1} q(e_1)\log P(e_1 | G) + \sum_{e_2} q(e_2)\log P(e_2 | G)+...
\end{eqnarray*}
The first gives a \emph{different} distribution over $\{e_i\}$ than EM:
\begin{eqnarray*}
\log q(e_i) &\stackrel{+}{=}& \sum_G q(G)\sum_{e_{j\not= i}} q(\{e_j\}) \log P(e_i, t_i, \{e_{j\not= i}\}, \{t_{j\not= i}\}, G) \\
&\stackrel{+}{=}& \sum_G q(G) q(e_i) \log P(e_i, t_i, G) \\
&\stackrel{+}{=}& \log P(t_i | e_i) + \sum_G q(G) \log P(e_i | G) \\
q(e_i) &\propto& P(t_i | e_i) \prod_G \left[ P(e_i | E)\right]^{q(G)}
\end{eqnarray*}
This can be approximated by weighting samples from the obvious proposal distribution
$$
r(e_i) = \sum_G P(e_i | G) q(G)
$$
A reasonable form for $q(G)$ is (yet another) factored distribution:
\begin{eqnarray*}
q(G) &=& \prod_k q(z_k)\\
q(z_k) &=& \mbox{Beta}(\alpha_k, \beta_k)
\end{eqnarray*}
where each $z_k$ is the $k$th production probability.
So, we could do gradient descent on  $\{\alpha_k\}\cup\{\beta_k\}$.
For computing description length, we could say that there is a Bernoulli RV for each production controlling whether it is even in the grammar, with a Beta prior.
Then, we can easily find the expected description length of a grammar drawn from $q(G)$.
\end{document}