### 2013-08-13 drafts/enumeration/enumeration.tex
Added writeup describing memory-efficient enumeration

\documentclass{article}

\begin{document}

The EC algorithm requires some means of obtaining many unique programs that have high likelihood under a given grammar.
We could do this by sampling from the grammar, but doing so can often fail to reveal shared substructures until a very large number of programs have been sampled.
A breadth-first enumeration tends to reveal more common substructure, but has memory consumption that grows exponentially with the number of enumerated programs.
This document describes an algorithm for performing memory-efficient (approximate) enumeration.

The general idea is to use the provided grammar to build a binary prefix-free code for parse trees, satisfying
$$
- \log P(z|G) \approx \ell (z|G)
$$
where $z$ is the parse tree, $G$ is the grammar, and $\ell(\cdot|\cdot)$ is conditional description length.
Then, we enumerate binary strings in lexicographic order and attempt to decode these strings in to parse trees.
Upon successful decoding of a binary string, the corresponding program is added to the frontier.
So, the space complexity of this algorithm is very low: we need to maintain one binary string (whose length grows logarithmically with the size of the frontier) as well as the set of programs we've already enumerated.
This algorithm can be seen as a form of iterative deepening, and so should have the same time complexity as the breadth-first enumeration algorithm.

The encoding of a parse tree is as follows.
First, a Huffman encoding tree is constructed from all available productions at the given choice point; that is, all productions whose type unifies with the current requested type.
Then, the chosen production is encoded using this tree.
Finally, the encodings for all non-terminals in this production are sequentially appended.

This encoding is best illustrated using an example.
Say that our library has only one terminal, $I$, having type $\alpha\to\alpha$.
The requested type for a given program is $\alpha\to\alpha$; suppose that the chosen program is $(I\; I)$ and that the grammar has probability 1/2 of choosing either $I$ or an application.

There's only one possible parse of $(I I)$.
First, an application is chosen.
So, construct a Huffman encoding tree for the two productions $S\to S S$ and $S\to I$; say that the code word for $S\to SS$ is 0 and the code for $S\to I$ is 1.
Then, the first bit is a 0 because the first production used in the parse is that of an application.
The encoding of $I$ is a 1, so the next two bits are both 1's, giving an encoding of $011$ for $(I\; I)$.

\end{document}