### 2014-06-08 old/drafts/ijcai2013/ResponsesToReviewers.txt
Moved everything into an a new 'old' directory.

Draft responses to IJCAI 2013 Review

overall common criticisms that should be addressed: 
1) is the work original (1, 4)?
2) how is the work related to ILP and predicate invention (1)
3) how is the work related to Chaitin's Algorithmic Complexity Theory (4)
4)

############################## Response to reviewer 1 ##############################
Main issue: What is the relationship to ILP?

Response: The reviewer is right to raise a connection between our work
and predicate invention in ILP, and we will revise the paper to note
this. However, while predicate invention in ILP is motivated by some
of the same considerations which motivate our work, the
learning framework in ILP is quite different from ours. ILP concerns
learning FOL theories from positive and negative
examples; our framework learns functional programs with respect to a
blackbox evaluator. This results in two main
differences. First, our framework is easily applicable to
both traditional ILP problems (like the boolean circuit learning
example) and problems more commonly found in statistical learning or
genetic programming (like the symbolic regression example). The second
and more crucial difference is that our choice of a functional
programming language makes it natural to capture reusable concepts as
higher-order functions (see, for example, the learned concept $E_1$ or
$E_2$ in Figure 4).

############################## Response to reviewer 2 ##############################
Main issue: Why not do a proper bayesian thing? Statistical discovery?

Response: The reviewer wonders about the algorithm's scalability and
points out that a sufficient number of solutions need to be found by
brute force search for the algorithm to work: we agree with this and
wish to emphasize that it is precisely the presence of some easy
problems that enables the algorithm to initially bootstrap an
effective grammar.

We agree that our approach can be framed as a statistical discovery
technique in which we iteratively re-estimate a stochastic grammar. We
are currently investigating this and will revise the paper to note
this interpretation. However, generating directly from this grammar,
as the reviewer suggests, instead of enumerating in a best-first
manner exacerbates the "needle-in-a-haystack" problem that our
algorithm faces; by enumerating programs, we are able to exhaustively
cover a large part of the prior mass of our stochastic grammar. We are
working to determine the advantages and disadvantages of
these two approaches.

############################## Response to reviewer 3 ##############################
Main issue: Why not use mdl measure like -log(p) instead of counting nodes.

Response: We would like to respond particularly to the reviewer's
question of why an mdl measure such as -log p was not used instead of
counting the nodes of the trees in Section 3.4. We agree with the
reviewer that this approach would be more elegant, and we will revise
the paper to make note of this. In order to take into account the
possibility of new rules (i.e. concept learning), however, we would
need to place a prior distribution over grammars (see, also, the
response to reviewer 2). This would likely require an expensive and
potentially inaccurate Monte Carlo approximation. We found the proxy
calculation of counting the number of unique subtrees in a set of
solutions to be a simple, intuitive, and elegant computation that gets
at the compressibility of the data.

############################## Response to reviewer 4 ##############################
Main issues: 1) Does this extend to noisy data. 2) What about Chaitin? (refer to response to reviewer 1).

Response: We understand the reviewer's concerns that a) this work has
similarities to Chaitin's Algorithmic Information Theory framework and
that b) that it is not clear in the paper how to extend this work to
noisy data. The reviewer is right that our approach is similar in
spirit to some of Chaitin's theoretical work, and his work will be
cited prominently, but our emphasis here is a novel search algorithm
that allows us to quickly bootstrap hundreds of programs for many
related tasks.

Dealing with noisy data requires one to choose a function that scores
candidate solutions to a task. We emphasize the aspects of our
approach that are independent of this, but there are straightforward
ways to extend our algorithm to noisy data. The most way natural is to
weight the subtree counting in section 3.4 by the chosen scoring
function. The paper will be revised to reflect this.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
TODO: implement josh's notes on this:

Notes from Josh: "We will revise the paper to note the The reviewers
are right to raise a connection to ILP and we will revise the paper to
note this, and we will revise the paper to note this, but we wills
tress the following differents.

Some of the same considerations not all
First order logic vs FOL.

We agree with the reviewer and we are currently investigating
(stochastic model thing) maybe move the we are going to do this part
to the start.

We will focus on this initially.

Typo "we would like to respond"

Can we argue about how the number of subtrees relating to the
nonparametric distribution. We will revise the paper to discuss this.
We do see what we are doing as an approximation to all that stuff, and
we will revise the paper to reflect that.

"It was not clear in our paper" instead of "It is not clear." and we
will revise the paper to reflect this. For the noisy data part.

Chaitin, should be cited prominently in the paper and we will.

"We believe there are straightforward ways" --> there are
straightforward ways and we are currently exploring them, and we will
revise the paper to reflect this.