### 2014-06-08 old/drafts/ObjectDetection/ObjectDetection.tex

\documentclass{article}

\title{Learning compositional object detectors}
\author{Eyal Dechter}

\begin{document}
\maketitle
\date
\section{Introduction}

At that heart of modern AI and machine learning is a tension between the
structured approaches to knowledge representation that dominate in classical
AI and bottom up data driven statistical techniques that deal flexibly with
large amounts of noisy and incomplete data. The field of computer vision is an
excellent example of this tension: on the one hand, visual data is extremely
high-dimensional, ambiguous, and noisy; on the other hand, the knowledge that
computer vision systems attempt to extract from images is highly structured.
The world behind the images is composed of objects and environments in
relations to one another, these objects are composed of other objects, and
both objects and their environments must obey the physical laws.

Recently, several attempts to bridge this gap have been proposed in the
subfield of computer vision that is concerned with object detection. One of
these approaches, due to Fleizen... and McAllester, works by constructing a
grammar over object detectors and learning from data the object detector that
best suits a given classification problem over object types. A detector drawn
from this grammar has at its leaves primitive feature detectors
(Fleizensomething et al. specifically used HOG detectors). The score of a
detector on an image is calculated recursively given the primitive feature
detectors.

Object detectors, which are drawn from a grammar, are a natural way to
describe the ways in which parts give rise to wholes in images of objects.

\end{document}