### 2014-06-08 old/drafts/aistats2013/abstracts.tex

\documentclass[twoside]{article}
\usepackage{aistats2014}

\begin{document}

\twocolumn[

\aistatstitle{Symbolic Density Estimation: Uncovering the Code.}

\aistatsauthor{ Anonymous Author 1 \And Anonymous Author 2 \And Anonymous Author 3 }

\aistatsaddress{ Unknown Institution 1 \And Unknown Institution 2 \And Unknown Institution 3 } ]

\begin{abstract}
One prominent view in Machine Learning is that to make accurate predictions and generalizations from data, we need to learn the latent probabilistic generative models that give rise to the data we observe. Graphical models and neural networks

\end{abstract}

\begin{abstract}

\end{abstract}

\end{document}